---
title: "ENVI 504 Data Project: Bacterial Growth Curves"
author: "Anna Vietmeier"
date: "7/26/2021"
output: html_document
---
---

## Introduction
Abandoned mine drainage (AMD) is the result of abandoned mining operations. AMD is high in sulfuric acid, heavy metals, and metalloids. Treatment of AMD can be carried out via passive remediation systems. Passive remediation systems consist of large settling ponds that promote the biogeochemical precipitation of AMD contaminants. Since these systems are open to the environment, they are naturally colonized by microbial populations. Microbes within these populations metabolically impact the remediation process, both positively and negatively. 
Isolate AV1 from the Boyce park passive remediation system pond 1 has been identified to positively impact iron remediation and negatively impact manganese remediation. Characterization of this microbial isolate is important to understand more about this novel isolate and understand how it can impact AMD remediation. Characterization of isolate AV1 includes understanding how the isolate grows under different conditions. This is facilitated via growing AV1 in rich media such as R2A as well as minimal media M63. 

## Methods
Isolate AV1 was grown in the shaker at 30C with either R2A pH 4.0 or M63 media. M63 minimal media was supplemented with an additional carbon source, either lactate or acetate. All conditions were grown at 30C. Different subculture ratios were used for the rich media to mirror those that had been manually obtained in the lab. Data for this experiment was collected using a 96-well plate and plate reader. Controls for the experiments included sterile R2A pH 4.0 and M63 supplemented with either lactate or acetate to be used as a blank and remove background noise. 

## Results
All <b> Figure A's</b> show each individual growth curve for the given condition, in total, each condition was tested 8 times. All <b> Figure B's</b> show the average and standard deviation of the 8 replicates. All <b>Figure C's</b> show the average exponential phase of growth for the replicates and also show an exponential trend line to determine the doubling time. 

```{r include = FALSE}

library(readr) 
library(dplyr)
library(ggplot2)
library(purrr)
library(matrixStats)
library(reshape2)
library(ggpubr)
library(broom) 
library(lattice)
library(Hmisc)
library(ggpmisc)
library(ggpubr)
library(gt)


setwd("~/Documents/Duquesne/Year_2_F20_S21_U21/Summer 2021/ENVI 404:504 Computer Tools for Scientists Excel & R/RStudio/ENVI504_6.24.21")

gc <- read_csv("AV_85.columns.csv", locale = locale(encoding = "Latin1"), skip = 17, col_names = TRUE) #load in datafile. Skip 1st 17 rows. Column names listed
```


### Growth Condition 1: AV1 R2A pH 4.0 30C 1:25 subculture 
<br>

```{r echo=FALSE}

#### Condition 1 Individual Replicates
eight1 <- ggplot(gc, aes(x = Time)) + 
  geom_point(aes(y=(A1-A8), color="A1")) +
  geom_point(aes(y=(B1-B8), color="B1")) +
  geom_point(aes(y=(C1-C8), color="C1")) +
  geom_point(aes(y=(D1-D8), color="D1")) +
  geom_point(aes(y=(E1-E8), color="E1")) +
  geom_point(aes(y=(F1-F8), color="F1")) +
  geom_point(aes(y=(G1-G8), color="G1")) +
  geom_point(aes(y=(H1-H8), color="H1")) +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  labs(x = "Hours", y = "OD600 (Log 10)") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),  ) +
  ggtitle("Individual Replicates") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.text = element_text(face = "plain", size = 12)) +
  theme(legend.position = "bottom") +
  theme(axis.title = element_text(size = 15)) 
print(eight1)
```

<b> Figure 1A. </b> Shows the growth curve experiment for condition 1 for 8-individual replicates.



```{r echo=FALSE}
### Condition 1 Average & Standard Deviation of Replicates
gc_1 <- gc %>%
  rowwise() %>%
  mutate(
    m.R2A = (mean(c(A1,B1,C1,D1,E1,F1,G1,H1)) - mean(c(A8,B8,C8,D8,E8,F8,G8,H8))),
    sd = sd(c(A1,B1,C1,D1,E1,F1,G1,H1))
    )


avg1 <- ggplot(gc_1, aes(x = Time, y = m.R2A)) + geom_point(alpha=0.7, color = "blue") +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  scale_y_log10() +
  ggtitle("Average of Replicates") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
    ) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  geom_errorbar(aes(ymin=m.R2A-sd, ymax=m.R2A+sd), width=.1,
                position=position_dodge(.9))
print(avg1)

```

<b> Figure 1B. </b> Shows the growth curve experiment for condition 1 as the average of 8-individual replicates and the standard deviation of the replicates.


```{r echo=FALSE}
### Condition 1 Exponential Growth
ex1 <- gc_1[c(5:26),c(1,99)]
x <- ex1$Time
y <- ex1$m.R2A
df <- data.frame(x, y)
exp.mod    <- lm(log(y) ~ x, df)
new_x      <- seq(min(x), max(x), 0.01)
prediction <- exp(predict(exp.mod, newdata = list(x = new_x)))
exp_line   <- data.frame(x = new_x, y = prediction)
ex1_plot <- ggplot(data = df, mapping = aes(x, y)) + 
  geom_point(alpha=0.7, color = "blue") +
  scale_y_log10() +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  ggtitle("Exponential Growth Phase of Replicates") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  geom_line(data = exp_line)

R_sq <- round(summary(exp.mod)$r.squared, 2)
B <- as.numeric(round(exp.mod$coefficients[2], 2))
dt <- log(2)/B

print(ex1_plot)

```

<b> Figure 1C. </b> Shows the exponential phase of growth from the growth curve experiment for condition 1 as the average of 8-individual replicates. An exponential trend line has been fitted to the data. 


## Growth Condition 2: AV1 in R2A pH 4.0 30C 1:50 subculture

```{r echo=FALSE}
###Condition 2 Individual Replicates
a2 <- ggplot(gc, aes(x = Time)) + 
  geom_point(aes(y=(A2-A8), color="A2")) +
  geom_point(aes(y=(B2-A8), color="B2")) +
  geom_point(aes(y=(C2-C8), color="C2")) +
  geom_point(aes(y=(D2-D8), color="D2")) +
  geom_point(aes(y=(E2-E8), color="E2")) +
  geom_point(aes(y=(F2-F8), color="F2")) +
  geom_point(aes(y=(G2-G8), color="G2")) +
  geom_point(aes(y=(H2-H8), color="H2")) +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  labs(x = "Hours", y = "OD600 (Log 10)") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  ggtitle("Individual Replicates") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  theme(legend.position = "bottom") +
  theme(axis.title = element_text(size = 15)) 
print(a2)
```

<b> Figure 2A. </b> Shows the growth curve experiment for condition 1 for 8-individual replicates.

```{r echo=FALSE}
### Condition 2 Average & Standard Deviation of Replicates
gc_2 <- gc %>%
  rowwise() %>%
  mutate(
    m2.R2A = (mean(c(A2,B2,C2,D2,E2,F2,G2,H2)) - mean(c(A8,B8,C8,D8,E8,F8,G8,H8))),
    sd = sd(c(A2,B2,C2,D2,E2,F2,G2,H2))
  )

ggplot(gc_2, aes(x = Time, y = m2.R2A)) + geom_point(alpha=0.7, color = "blue") +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  scale_y_log10() +
  ggtitle("AV1 R2A pH 4.0 30C 1:50 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  geom_errorbar(aes(ymin=m2.R2A-sd, ymax=m2.R2A+sd), width=.1,
                position=position_dodge(.9))
```

<b> Figure 2B. </b> Shows the growth curve experiment for condition 1 as the average of 8-individual replicates and the standard deviation of the replicates.

```{r echo=FALSE}
### Condition 2 Exponential Growth
ex2 <- gc_2[c(5:26),c(1,99)]
x.2 <- ex2$Time
y.2 <- ex2$m2.R2A

df2 <- data.frame(x.2, y.2)

exp.mod2    <- lm(log(y.2) ~ x.2, df2)

new_x2      <- seq(min(x.2), max(x.2), 0.01)
prediction2 <- exp(predict(exp.mod2, newdata = list(x.2 = new_x2)))
exp_line2   <- data.frame(x.2 = new_x2, y.2 = prediction2)

ex2_plot <- ggplot(data = df2, mapping = aes(x.2, y.2)) + 
  geom_point(alpha=0.7, color = "blue") +
  scale_y_log10() +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  ggtitle("Exponential Phase AV1 R2A pH 4.0 30C 1:50 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  geom_line(data = exp_line2)

ex2_plot
R_sq.2 <- round(summary(exp.mod2)$r.squared, 2)
B.2 <- as.numeric(round(exp.mod2$coefficients[2], 2))
dt.2 <- log(2)/B.2

```

<b> Figure 2C. </b> Shows the exponential phase of growth from the growth curve experiment for condition 1 as the average of 8-individual replicates. An exponential trend line has been fitted to the data. 

## Growth Condition 3: AV1 R2A pH 4.0 30C 1:100 subculture

```{r echo=FALSE, warning=FALSE}
###Condition 3 Individual Replicates
a3 <- ggplot(gc, aes(x = Time)) + 
  geom_point(aes(y=A3-A8, color="A3")) +
  geom_point(aes(y=B3-B8, color="B3")) +
  geom_point(aes(y=C3-C8, color="C3")) +
  geom_point(aes(y=D3-D8, color="D3")) +
  geom_point(aes(y=E3-E8, color="E3")) +
  geom_point(aes(y=F3-F8, color="F3")) +
  geom_point(aes(y=G3-G8, color="G3")) +
  geom_point(aes(y=H3-H8, color="H3")) +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  labs(x = "Hours", y = "OD600 (Log 10)") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  ggtitle("Individual Replicates") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  theme(legend.position = "bottom") +
  theme(axis.title = element_text(size = 15)) 
print(a3)
```

<b> Figure 3A. </b> Shows the growth curve experiment for condition 1 for 8-individual replicates.

```{r echo=FALSE, warning=FALSE}
### Condition 3 Average & Standard Deviation of Replicates
gc_3.R2A <- gc %>%
  rowwise() %>%
  mutate(
    m3.R2A = (mean(c(A3,B3,C3,D3,E3,F3,G3,H3 - mean(c(A8,B8,C8,D8,E8,F8,G8,H8))))),
    sd3.R2A = sd(c(A3,B3,C3,D3,E3,F3,G3,H3))
  )

ggplot(gc_3.R2A, aes(x = Time, y = m3.R2A)) + geom_point(alpha=0.7, color = "blue") +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  scale_y_log10() +
  ggtitle("AV1 R2A pH 4.0 30C 1:100 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  geom_errorbar(aes(ymin=m3.R2A-sd3.R2A, ymax=m3.R2A+sd3.R2A), width=.1,
                position=position_dodge(.9))
```

<b> Figure 3B. </b> Shows the growth curve experiment for condition 1 as the average of 8-individual replicates and the standard deviation of the replicates.

```{r echo=FALSE, warning=FALSE}
### Condition 3 Exponential Growth
ex3 <- gc_3.R2A[c(9:36),c(1,99)]
x.3 <- ex3$Time
y.3 <- ex3$m3.R2A
df3 <- data.frame(x.3, y.3)
exp.mod3    <- lm(log(y.3) ~ x.3, df3)
new_x3      <- seq(min(x.3), max(x.3), 0.01)
prediction3 <- exp(predict(exp.mod3, newdata = list(x.3 = new_x3)))
exp_line3   <- data.frame(x.3 = new_x3, y.3 = prediction3)

ex3_plot <- ggplot(data = df3, mapping = aes(x.3, y.3)) + 
  geom_point(alpha=0.7, color = "blue") +
  scale_y_log10() +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  ggtitle("Exponential Phase AV1 R2A pH 4.0 30C 1:100 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  geom_line(data = exp_line3)
ex3_plot
R_sq.3 <- round(summary(exp.mod3)$r.squared, 2)
B.3 <- as.numeric(round(exp.mod3$coefficients[2], 2))
dt.3 <- log(2)/B.3
```

<b> Figure 3C. </b> Shows the exponential phase of growth from the growth curve experiment for condition 1 as the average of 8-individual replicates. An exponential trend line has been fitted to the data. 

## Growth Condition 4: AV1 R2A pH 4.0 30C 1:200 subculture

```{r echo=FALSE, warning=FALSE}
###Condition 4 Individual Replicates
a4 <- ggplot(gc, aes(x = Time)) + 
  geom_point(aes(y=A4-A8, color="A4")) +
  geom_point(aes(y=B4-B8, color="B4")) +
  geom_point(aes(y=C4-C8, color="C4")) +
  geom_point(aes(y=D4-D8, color="D4")) +
  geom_point(aes(y=E4-E8, color="E4")) +
  geom_point(aes(y=F4-F8, color="F4")) +
  geom_point(aes(y=G4-G8, color="G4")) +
  geom_point(aes(y=H4-H8, color="H4")) +
  ggtitle("Individual Replicates") +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  labs(x = "Hours", y = "OD600 (Log 10)") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  theme(legend.position = "bottom") +
  theme(axis.title = element_text(size = 15)) 
print(a4)
```

<b> Figure 4A. </b> Shows the growth curve experiment for condition 1 for 8-individual replicates.


```{r echo=FALSE}
### Condition 4 Average & Standard Deviation of Replicates
gc_4.R2A <- gc %>%
  rowwise() %>%
  mutate(
    m4.R2A = (mean(c(A4,B4,C4,D4,E4,F4,G4,H4)) - - mean(c(A8,B8,C8,D8,E8,F8,G8,H8))),
    sd = sd(c(A4,B4,C4,D4,E4,F4,G4,H4))
  )

ggplot(gc_4.R2A, aes(x = Time, y = m4.R2A)) + geom_point(alpha=0.7, color = "blue") +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  scale_y_log10() +
  ggtitle("AV1 R2A pH 4.0 30C 1:200 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  geom_errorbar(aes(ymin=m4.R2A-sd, ymax=m4.R2A+sd), width=.1,
                position=position_dodge(.9))
```

<b> Figure 4B. </b> Shows the growth curve experiment for condition 1 as the average of 8-individual replicates and the standard deviation of the replicates.

```{r echo=FALSE}
### Condition 4 Exponential Growth
ex4 <- gc_4.R2A[c(12:36),c(1,99)]
x.4 <- ex4$Time
y.4 <- ex4$m4.R2A
df4 <- data.frame(x.4, y.4)
exp.mod4    <- lm(log(y.4) ~ x.4, df4)
new_x4      <- seq(min(x.4), max(x.4), 0.01)
prediction4 <- exp(predict(exp.mod4, newdata = list(x.4 = new_x4)))
exp_line4   <- data.frame(x.4 = new_x4, y.4 = prediction4)
ex4_plot <- ggplot(data = df4, mapping = aes(x.4, y.4)) + 
  geom_point(alpha=0.7, color = "blue") +
  scale_y_log10() +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  ggtitle("Exponential Phase AV1 R2A pH 4.0 30C 1:200 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  geom_line(data = exp_line4)
ex4_plot
R_sq.4 <- round(summary(exp.mod4)$r.squared, 2)
B.4 <- as.numeric(round(exp.mod4$coefficients[2], 2))
dt.4 <- log(2)/B.4
```

<b> Figure 4C. </b> Shows the exponential phase of growth from the growth curve experiment for condition 1 as the average of 8-individual replicates. An exponential trend line has been fitted to the data. 

## Growth Condition 5: AV1 R2A pH 4.0 30C 1:400 subculture

```{r echo=FALSE, warning=FALSE}
###Condition 5 Individual Replicates
a5 <- ggplot(gc, aes(x = Time)) + 
  geom_point(aes(y=A5-A8, color="A5")) +
  geom_point(aes(y=B5-B8, color="B5")) +
  geom_point(aes(y=C5-C8, color="C5")) +
  geom_point(aes(y=D5-D8, color="D5")) +
  geom_point(aes(y=E5-E8, color="E5")) +
  geom_point(aes(y=F5-F8, color="F5")) +
  geom_point(aes(y=G5-G8, color="G5")) +
  geom_point(aes(y=H5-H8, color="H5")) +
  ggtitle("Individual Replicates") +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  labs(x = "Hours", y = "OD600 (Log 10)") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  theme(legend.position = "bottom") +
  theme(axis.title = element_text(size = 15)) 
print(a5)
```

<b> Figure 5A. </b> Shows the growth curve experiment for condition 1 for 8-individual replicates.

```{r echo=FALSE, warning=FALSE}
### Condition 5 Average & Standard Deviation of Replicates
gc_5.R2A <- gc %>%
  rowwise() %>%
  mutate(
    m5 = (mean(c(A5,B5,C5,D5,E5,F5,G5,H5)) - mean(c(A8,B8,C8,D8,E8,F8,G8,H8))),
    sd = sd(c(A5,B5,C5,D5,E5,F5,G5,H5))
  )
ggplot(gc_5.R2A, aes(x = Time, y = m5)) + geom_point(alpha=0.7, color = "blue") +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  scale_y_log10() +
  ggtitle("AV1 R2A pH 4.0 30C 1:400 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  geom_errorbar(aes(ymin=m5-sd, ymax=m5+sd), width=.1,
                position=position_dodge(.9))
```

<b> Figure 5B. </b> Shows the growth curve experiment for condition 1 as the average of 8-individual replicates and the standard deviation of the replicates.

```{r echo=FALSE}
### Condition 5 Exponential Growth
ex5 <- gc_5.R2A[c(5:26),c(1,99)]
x.5 <- ex5$Time
y.5 <- ex5$m5

df5 <- data.frame(x.5, y.5)

exp.mod5    <- lm(log(y.5) ~ x.5, df5)

new_x5      <- seq(min(x.5), max(x.5), 0.01)
prediction5 <- exp(predict(exp.mod5, newdata = list(x.5 = new_x5)))
exp_line5   <- data.frame(x.5 = new_x5, y.5 = prediction5)

ex5_plot <- ggplot(data = df5, mapping = aes(x.5, y.5)) + 
  geom_point(alpha=0.7, color = "blue") +
  scale_y_log10() +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  ggtitle("Exponential Phase AV1 R2A pH 4.0 30C 1:400 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  geom_line(data = exp_line5)
ex5_plot
R_sq.5 <- round(summary(exp.mod5)$r.squared, 2)
B.5 <- as.numeric(round(exp.mod5$coefficients[2], 2))
dt.5 <- log(2)/B.5
```

<b> Figure 5C. </b> Shows the exponential phase of growth from the growth curve experiment for condition 1 as the average of 8-individual replicates. An exponential trend line has been fitted to the data. 

## Growth Condition 6: AV1 in M63+L 30C 1:10 subculture

```{r echo=FALSE, warning=FALSE}
###Condition 6 Individual Replicates
a6 <- ggplot(gc, aes(x = Time)) + 
  geom_point(aes(y=A6-A9, color="A6")) +
  geom_point(aes(y=B6-B9, color="B6")) +
  geom_point(aes(y=C6-C9, color="C6")) +
  geom_point(aes(y=D6-D9, color="D6")) +
  geom_point(aes(y=E6-E9, color="E6")) +
  geom_point(aes(y=F6-F9, color="F6")) +
  geom_point(aes(y=G6-G9, color="G6")) +
  geom_point(aes(y=H6-H9, color="H6")) +
  ggtitle("Individual Replicates") +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  labs(x = "Hours", y = "OD600 (Log 10)") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  theme(legend.position = "bottom") +
  theme(axis.title = element_text(size = 15)) 
print(a6)
```

<b> Figure 6A. </b> Shows the growth curve experiment for condition 1 for 8-individual replicates.

```{r echo=FALSE}
### Condition 6 Average & Standard Deviation of Replicates
gc_6.M <- gc %>%
  rowwise() %>%
  mutate(
    m6 = (mean(c(A6,B6,C6,D6,E6,F6,G6,H6)) - mean(c(A9,B9,C9,D9,E9,F9,G9,H9))),
    sd = sd(c(A6,B6,C6,D6,E6,F6,G6,H6))
  )
ggplot(gc_6.M, aes(x = Time, y = m6)) + geom_point(alpha=0.7, color = "blue") +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  scale_y_log10() +
  ggtitle("AV1 in M63+L 30C 1:10 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  geom_errorbar(aes(ymin=m6-sd, ymax=m6+sd), width=.1,
                position=position_dodge(.9))
```

<b> Figure 6B. </b> Shows the growth curve experiment for condition 1 as the average of 8-individual replicates and the standard deviation of the replicates.


```{r echo=FALSE}
### Condition 6 Exponential Growth
ex6 <- gc_6.M[c(6:80),c(1,99)]
x.6 <- ex6$Time
y.6 <- ex6$m6

df6 <- data.frame(x.6, y.6)

exp.mod6    <- lm(log(y.6) ~ x.6, df6)

new_x6      <- seq(min(x.6), max(x.6), 0.01)
prediction6 <- exp(predict(exp.mod6, newdata = list(x.6 = new_x6)))
exp_line6   <- data.frame(x.6 = new_x6, y.6 = prediction6)

ex6_plot <- ggplot(data = df6, mapping = aes(x.6, y.6)) + 
  geom_point(alpha=0.7, color = "blue") +
  scale_y_log10() +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  ggtitle("Exponential Phase AV1 in M63+L 30C 1:10 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  geom_line(data = exp_line6)
print(ex6_plot)
R_sq.6 <- round(summary(exp.mod6)$r.squared, 2)
B.6 <- as.numeric(round(exp.mod6$coefficients[2], 2))
dt.6 <- log(2)/B.6

```

<b> Figure 6C. </b> Shows the exponential phase of growth from the growth curve experiment for condition 1 as the average of 8-individual replicates. An exponential trend line has been fitted to the data. 

## Growth Condition 7: AV1 in M63+A 30C 1:10 subculture

```{r echo=FALSE, warning=FALSE}
###Condition 7 Individual Replicates
a7 <- ggplot(gc, aes(x = Time)) + 
  geom_point(aes(y=A7-A10, color="A7")) +
  geom_point(aes(y=B7-B10, color="B7")) +
  geom_point(aes(y=C7-C10, color="C7")) +
  geom_point(aes(y=D7-D10, color="D7")) +
  geom_point(aes(y=E7-E10, color="E7")) +
  geom_point(aes(y=F7-F10, color="F7")) +
  geom_point(aes(y=G7-G10, color="G7")) +
  geom_point(aes(y=H7-H10, color="H7")) +
  ggtitle("Individual Replicates") +
  scale_y_log10() +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  labs(x = "Hours", y = "OD600 (Log 10)") +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  theme(legend.position = "bottom") +
  theme(axis.title = element_text(size = 15)) 
print(a7)
```

<b> Figure 7A. </b> Shows the growth curve experiment for condition 1 for 8-individual replicates.

```{r echo=FALSE}
### Condition 7 Average & Standard Deviation of Replicates
gc_7M <- gc %>%
  rowwise() %>%
  mutate(
    m7 = (mean(c(A7,B7,C7,D7,E7,F7,G7,H7)) - mean(c(A10,B10,C10,D10,E10,F10,G10,H10))),
    sd = sd(c(A7,B7,C7,D7,E7,F7,G7,H7))
  )
ggplot(gc_7M, aes(x = Time, y = m7)) + geom_point(alpha=0.7, color = "blue") +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  scale_x_continuous(breaks = seq(0, 28, by = 2)) +
  scale_y_log10() +
  ggtitle("AV1 in M63+A 30C 1:10 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  theme(axis.text = element_text(face = "plain", size = 12)) + 
  geom_errorbar(aes(ymin=m7-sd, ymax=m7+sd), width=.1,
                position=position_dodge(.9))
```

<b> Figure 7B. </b> Shows the growth curve experiment for condition 1 as the average of 8-individual replicates and the standard deviation of the replicates.

```{r echo=FALSE}
### Condition 7 Exponential Growth
ex7 <- gc_7M[c(5:90),c(1,99)]
x.7 <- ex7$Time
y.7 <- ex7$m7
df7 <- data.frame(x.7, y.7)
exp.mod7    <- lm(log(y.7) ~ x.7, df7)
new_x7      <- seq(min(x.7), max(x.7), 0.01)
prediction7 <- exp(predict(exp.mod7, newdata = list(x.7 = new_x7)))
exp_line7   <- data.frame(x.7 = new_x7, y.7 = prediction7)

ex7_plot <- ggplot(data = df7, mapping = aes(x.7, y.7)) + 
  geom_point(alpha=0.7, color = "blue") +
  scale_y_log10() +
  labs(x = "Time (Hours)", y = "OD600 (Log 10)") +
  theme(axis.title = element_text(size = 15)) +
  ggtitle("Exponential Phase AV1 in M63+A 30C 1:10 subculture") +
  theme(plot.title = element_text(size = 16)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(panel.background = element_rect(fill = "white", colour = "black")) +
  theme(aspect.ratio = 1) +
  theme(
    panel.grid.major.x = element_line(color = "grey80", linetype = "dashed"),
    panel.grid.minor.x = element_line(colour = "grey85", linetype = "dashed"),
  ) +
  geom_line(data = exp_line7)
print(ex7_plot)
R_sq.7 <- round(summary(exp.mod7)$r.squared, 2)
B.7 <- as.numeric(round(exp.mod7$coefficients[2], 2))
dt.7 <- log(2)/B.7
```

<b> Figure 7C. </b> Shows the exponential phase of growth from the growth curve experiment for condition 1 as the average of 8-individual replicates. An exponential trend line has been fitted to the data. 



## Doubling Time for All Conditions 
```{r echo=FALSE}
Cons2 <- data.frame("Growth_Condition" =c("R2A pH 4.0 1:25", "R2A pH 4.0 1:50", "R2A pH 4.0 1:100", 
                                   "R2A pH 4.0 1:200", "R2A pH 4.0 1:400", "M63+Lactate 1:10", 
                                                   "M63+Acetate 1:10"),
                   "Doubling_Time" =c(dt, dt.2, dt.3, dt.4, dt.5, dt.6, dt.7),
                   "R_squared" = c(R_sq, R_sq.2, R_sq.3, R_sq.4, R_sq.5, R_sq.6, R_sq.7)
                   )

Cons2 %>%
  gt() %>%
  tab_header( title = "Isolate AV1", subtitle = "30C") %>%
  fmt_number(columns = vars(R_squared), decimals = 2) %>%
  fmt_number(columns = vars(Doubling_Time), decimals = 2) #changes that number of decimals in the table to the hundredths place


```



<b> Table 1.</b> Doubling times and R-squared values for the exponential phase of growth, calculated from the average of eight replicates for each condition. 

## Discussion
Bacterial growth curves of isolate AV1 was carried out in a 96-well plate and readings were taken at OD600 using a platereader. Controls for the experiment included the use of sterile media that would also double as a blank to remove background noise from the growth curves. All r-square values reported for the exponential trend line are within 0.04 to 1, this indicates that the line is a good fit for the data. Although, all R2A replicates were expected to have the same doubling time, this was not found to be true as indicated by table 1. There was an increased length of time required for the bacterial cells to double when inoculated into minimal media M63 supplemented with a carbon source, as expected. It also appears that isolate AV1 grows better when supplemented with acetate as a carbon source than lactate, which is also as expected, since lactate is a poor carbon source. 