---
title: "Final Homework"
author: "Anna Vietmeier"
date: "7/23/2021"
output: html_document
---

## Precipitation Patterns  
*Albasini Dam, South Africa*  

### Introduction 
**Due 27 July 2021** 

Analysis of data collected from the Albasini Dam weather station located in South Africa. Data consists of monthly precipitation measured in inches from 1980 to 2020. The hydrological year in South Africa runs from July to August. Data was converted to the hydrological year prior to beginning data analysis

The average precipitation for the hydrological year from 1980-2020 is represented in a bar plot (<b>Figure 1</b>). 
The average monthly precipitation from 1980-2020 and the standard deviation for each month is represented in a bar plot with error bars (<b>Figure 2</b>). Annual precipitation data from 1980-2000 and 2001-2020 were grouped together to compare for significant difference in perciptation using a t-test (<b>Figure 3</b>). 


```{r include=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(forcats)

# the data are provided via URL, the "filename" will not need adjustment - this may not apply to the options of the command.  If you prefer to work locally, the dataset is in the GitHub repository and the Box folder.  For the purposes of this homework, you may ignore the quality codes (in fact, you need to in order to process all of the data).  
prcp <- read_csv("https://duq.box.com/shared/static/d1523vk9rnilrtsuuhodffwj6vjygywk.csv", skip = 9)
prcp <- prcp[1:14610,]

# process dates with lubridate.  For help, see the lubridate cheat sheet in the Box folder.

###need to parse date: Year, Mon, Day
prcp$dt <- as_date(ymd(prcp$DATE))
```



```{r include=FALSE}
# Pull year and month from dt
prcp$year <- year(prcp$dt)
prcp$mont <- month(prcp$dt)

# Preallocate "hydro.y" for hydrologic year and hydro.m
prcp$hydro.y <-prcp$year
prcp$hydro.m <- prcp$mont
#  Use fn to add water year to the dataframe
# DEFINE: hydrologic year is the year that the hydrologic year ENDS.  Ex: Hydrologic year 2020 begins August 2019 and goes to July 2020.

########
# monthly data has issue. starts @ 6 and stops at month 12
   
for (i in 1:nrow(prcp)) {
      if (prcp$mont[i] >= 8) {
          prcp$hydro.y[i] <- prcp$hydro.y[i] + 1 
          prcp$hydro.m[i] <- prcp$hydro.m[i] - 7
      }

       else {
            prcp$hydro.m[i] <- prcp$hydro.m[i] + 5 
      }
}
# adds one year if the month is August or later.
# pulls month back so that August(8) is month 1
# pushes month forward so January(1) is month 6
```


### Annual Precipitation

```{r echo=FALSE}



annual <- prcp %>%
       group_by(hydro.y)%>%
       summarize(total = sum(`DAILY RAIN`, na.rm = TRUE)) 

ggplot(annual, aes(x = hydro.y, y = total)) +
      geom_bar(stat = "identity", fill = "steelblue") +
      labs(x= "Year", y = "Precipitation (inches)") +
      theme(panel.background = element_rect(fill = "white", colour = "black")) +
      theme(aspect.ratio = 1) +
      ggtitle("Annual Precipitation") +
      theme(axis.text = element_text(face = "plain", size = 12))
```


<b>Figure 1.</b> Albansi Dam, South Africa annual precipitation reported in inches from 1980-2020. The highest amount of precipitation was reported in 2000, and the lowest amount of precipitation was reported in 2005.

### Typical Monthly Precipitation

```{r echo=FALSE}
prcp$hydro.ym <- 100 * prcp$hydro.y + prcp$hydro.m


monthly <- prcp %>%
      group_by(hydro.ym) %>%
      summarize(total = sum(`DAILY RAIN`, na.rm = TRUE))



mon_pattern <- monthly %>%
      mutate(year = floor(hydro.ym/100)) %>%
      mutate(mon = hydro.ym - (100 * floor(hydro.ym/100))) %>%
      group_by(mon) %>%
      summarize(prcp = mean(total), std = sd(total))
mon_pattern$m <- factor(c("Aug", "Sep", "Oct", "Nov", "Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul")) # this puts the months in this order; otherwise, alphabetical is used.
ggplot(mon_pattern, aes(x = fct_reorder(m, mon), y = prcp)) + # fct_reorder is a command from forcats library that orders the factor, m (the name of the month)
      geom_bar(stat = "identity", fill = "steelblue") +
      geom_errorbar(aes(ymin=prcp-std, ymax=prcp+std), width=.1,
                position=position_dodge(.9)) +
      labs(x = "Month", y = "Average Monthly Precipitation (inches)") +
      theme(panel.background = element_rect(fill = "white", colour = "black")) +
      theme(aspect.ratio = 1) +
      ggtitle("Monthly Precipitation") +
      theme(axis.text = element_text(face = "plain", size = 12))
```


<b>Figure 2.</b> Albansi Dam, South Africa monthly average precipitation for the hydrological year from 1980-2020. Error bars depict standard deviation in precipitation for each month. The greatest amount of precipitation occured in Jan, followed by Feb. With the least amount of precipitation occuring in Jun, July, and August. The month of Feb experienced the largest fluxuation in precipitation as indicated by having the largest error bar. 

### T-test for the Scores 1980-2000 compared to 2001-2020

```{r echo=FALSE}
# t.test


prcp1980 <- prcp %>%
      filter(hydro.y < 2001) %>%
      group_by(hydro.y) %>%
      summarize(total = sum(`DAILY RAIN`, na.rm = TRUE))
prcp2000 <- prcp %>%
      filter(hydro.y > 2000) %>%
      group_by(hydro.y) %>%
      summarize(total = sum(`DAILY RAIN`, na.rm = TRUE))
t <- t.test(prcp1980,prcp2000)
annual$score <- as.character(annual$hydro.y) # preallocation
for (i in 1:nrow(annual)) {
      if (annual$score[i] < 2001) {
            annual$score[i] <- "1981-2000"
      } else {
            annual$score[i] <- "2001-2020"
      }
}



ggplot(annual, aes(x=score, y=total)) +
      geom_boxplot(outlier.colour = "red", fill='Steel Blue') +
      stat_summary(fun = "mean", geom = "point", shape = 8, size =2, color = 'Black') +
      labs(x = "Year", y = "Annual Precipitation (mm)") +
      theme(panel.background = element_rect(fill = "white", colour = "black")) +
      theme(aspect.ratio = 1) +
      ggtitle("Annual Precipitation 1981-2000 vs. 2001-2020") +
      theme(axis.text = element_text(face = "plain", size = 12))
```


<b>Figure 3.</b> Albansi Dam, South Africa comparison of annual precipitation from 1980-2000 to 2001-2020. Data has been grouped by the hydrological year for South Africa that starts in July. The mean of the data set is indicated by the astricks(*). Outliers in the dataset are indicated as a red dot. The t-test gave value p = 0.749, which is greater than 0.05, therefore this suggests the two populations are not statisticaly different. Although not statistically significant, the trend shown by the boxplot is consistant with the overall decrease in precipitation in the region.  


